---
title: "Data Sci 101 Project"
author: "Blue"
date: "2023-08-15"
output:
  word_document: default
  html_document: default
---





```{r}
set.seed(123); rnorm(1)
library(readr)
library(caret)
ames <- read_csv("ames2000_NAfix.csv") # read in data
library(ISLR)
HousingData = data.frame(ames) # turn into a dataframe





```

```{r}
#clean data


# Calculate the average of each numeric column
contVar <- c("Lot.Frontage", "Lot.Area", "Mas.Vnr.Area", "BsmtFin.SF.2", "BsmtFin.SF.2", "Bsmt.Unf.SF", "Total.Bsmt.SF", "X1st.Flr.SF", "X2nd.Flr.SF", "Enclosed.Porch", "X3Ssn.Porch", "Screen.Porch", "Pool.Area", "Misc.Val", "SalePrice") # collecting all continous variables to clean

for (i in contVar) {
  HousingData[, i] <- as.numeric(HousingData[, i])
}# turning each variable into a numeric

numeric_columns <- sapply(HousingData, is.numeric)

for (col_name in names(HousingData)[numeric_columns]) {
  avg_value <- mean(HousingData[[col_name]], na.rm = TRUE)
  HousingData[[col_name]][is.na(HousingData[[col_name]])] <- avg_value
} 


#for(i in 1:ncol(HousingData)){
  #if(is.numeric(HousingData[,i])){
  #  mean = mean(HousingData[,i], na.rm = T)}
  #for(i in 1:nrow(HousingData)){
 # if(is.na(HousingData[,i][x])){
 #  HousingData[,i][x] = mean}
#}}
  


for (i in 1:length(HousingData[[col_name]])) {
  if (is.na(HousingData[[col_name]][i])) {
    HousingData[[col_name]][i] <- average_value
  }# getting rid of a NA's and replacing with the average
}



```

```{r}
#clean catagorical data 



```




```{r}
#split data
smp_size = floor(.5 * nrow(HousingData))
train_Index = sample(seq_len(nrow(HousingData)), size = smp_size)
trainData = HousingData[train_Index, ]
testData = HousingData[-train_Index, ]

```
1) What field does the data come from?
This data comes from the Ames Assessor's Office.

• What are the goals of the study? Are there any effects of particular interest?
The goals of the study are to predict Salesprice based on the other given varaibles.

• How might these goals be answered, i.e. tests / confidence intervals?
This goal can be answered with linear regression taking into account all of the relevant variables.

```{r}
library(corrplot)


# Calculate correlation matrix
cor_matrix <- cor(trainData[, numeric_columns])
corrplot(cor_matrix)


```

2) How are the predictor variables spread out? Are there any noteworthy features to their spread
that could be highly influential observations?
• Are any of the predictor variables highly correlated?
Yes

```{r}
# Load ggplot2 library
library(ggplot2)
# boxplot of each variable
boxplot(trainData$Lot.Frontage, main = "Lot.Frontage")
boxplot(trainData$Lot.Area, main = "Lot.Area")
boxplot(trainData$Overall.Qual, main = "Overall.Qual")
boxplot(trainData$Overall.Cond, main = "Overall.Cond")
boxplot(trainData$Year.Built, main = "Year.Built")
boxplot(trainData$Year.Remod.Add, main = "Year.Remod.Add")
boxplot(trainData$TotRms.AbvGrd, main = "TotRms.AbvGrd")
boxplot(trainData$Full.Bath, main = "Full.Bath")
boxplot(trainData$Mas.Vnr.Area, main = "Mas.Vnr.Area")
boxplot(trainData$Total.Bsmt.SF, main = "Total.Bsmt.SF")
boxplot(trainData$X1st.Flr.SF, main = "X1st.Flr.SF")
boxplot(trainData$Garage.Area, main = "Garage.Area")
boxplot(trainData$Gr.Liv.Area, main = "Gr.Liv.Area")
boxplot(trainData$Fireplaces, main = "Fireplaces")
boxplot(trainData$Garage.Cars, main = "Garage.Cars")
#histogram of salePrice
hist(trainData$SalePrice)
median(trainData$SalePrice)
mean(trainData$SalePrice)
```


3) • Which predictor variables, if any, should be included in the model a priori?
Lot.Frontage, Lot.area, Overall.Qual, Year.Built, Year.Remod.Add, Mas.Vnr.Area, Total.Bsmt.SF, X1st.Flr.SF, Gl.Liv.Area, Full.Bath, TotRms.AbvGrd, Fireplaces, Garage.Cars, Garage.Area
• Are there any interactions that should be considered for inclusion in the model?
Yes, the Garage.Cars and Garage.Area should be considered since the two variables are highly related, if there are more garage car spots, the garage area will usually be bigger.
• Are there any three way interactions that should be considered?
 Overall.Qual, Overall.Cond, Year.Built, Year.Remod.Add, TotRms.AbvGrd, Full.Bath. This is a six way interaction but it should be considered since as each of these goes up, the saleprice will also go up.
• Are there any interactions that should NOT be considered?



```{r}
# Fit a linear regression model using other varaibles as the predictors
model <- lm(log(SalePrice) ~ Lot.Frontage * Lot.Area + Overall.Qual*Overall.Cond * Year.Built * Year.Remod.Add*Full.Bath*Half.Bath  + Mas.Vnr.Area + Total.Bsmt.SF * X1st.Flr.SF  +  Gr.Liv.Area * Fireplaces*  Garage.Cars * Garage.Area+TotRms.AbvGrd, data = trainData)

# Plot the model
plot(model)
summary(model)
# Calculate confidence intervals 
conf_interval <- confint(model)
print(conf_interval)




```

The residuals graph idicates that the linear regression assumptions (linear correlation, independence of residuals, normality of residuals)

QQ, the graph follows a diagonal line which is good since it means that the residuals are normaly distributed. At the very end, the points deviate however the vast majority of the points indicate the residuals are normally distributed.

the scale location plot is lightly curved out of being horizontal, this means that the variability of residuals is not consistent across different levels of the predictor variables.
```{r}
# Calculate the MSE
actual <- testData$SalePrice
predicted <- predict(model, newdata = testData)
mse <- mean((log(actual) - predicted))

# Print the calculated MSE
cat("Mean Squared Error (MSE):", mse, "\n") #log fits data better
 


```

```{r}
# Make predictions using the model and test data
predictions <- predict(model, newdata = testData)

# Plot the predicted values against the actual SalePrice in the test data
plot(testData$SalePrice, exp(predictions),
     xlab = "Actual SalePrice",
     ylab = "Predicted SalePrice",
     main = "Predicted vs. Actual SalePrice",
     xlim=c(0,800000),
     ylim=c(0, 800000)
     )
library(ggplot2)

#scatter plot of salePrice against livArea
ggplot(data = trainData, aes(x = SalePrice, y = Gr.Liv.Area)) +
  geom_point() +
  labs(x = "SalePrice", y = "Gr.Liv.Area", title = "Scatter Plot: SalePrice vs. Gr.Liv.Area")

# Calculate summary statistics for the predictions
summary(exp(predictions))


```

```{r}
# bototrap the mean
boot = function(x,stat, B=1000){
  n = length(x)
  boot_stats = matrix(nrow = B)
  boot_average = numeric(B)
  for(b in 1:1000){#B =1000
    boot_sample = sample(x,n, replace = TRUE)
    boot_stats[b] = mean(boot_sample)
    
  }
  return(boot_stats)}

SE = mean(boot(trainData$SalePrice, sd))


#95% confidence interval of the Means (SEM)
mean(exp(predictions))+1.96*SE # upper 97.5
median(exp(predictions))
mean(exp(predictions))
mean(exp(predictions))-1.96*SE # lower 2.5


```





https://stats.oarc.ucla.edu/r/modules/coding-for-categorical-variables-in-regression-models/
